\documentclass[a4paper,11pt]{letter}

\usepackage[left=3cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\signature{Dr. Denis Demidov}
\address{
Joint Supercomputer Center \\
Russian Academy of Sciences \\
Lobachevsky st. 2/31\\
420011 Kazan, Russia 
}
\begin{document}

\begin{letter}{
    Hans Petter Langtangen\\
    Editor-in-Chief\\
    SIAM Journal of Scientific Computing
    }
\opening{Dear Dr. Langtangen,}

On behalf of all the authors, I would like to ask you to consider paper
entitled ``Programming CUDA and OpenCL: a Case Study Using Modern C++
Libraries'' for publication in the ``Software and High-Performance Computing''
section of SIAM Journal of Scientific Computing.

In the paper we present a comparison of several modern C++ libraries providing
high-level interfaces for programming multi- and many-core architectures on top
of CUDA or OpenCL.  We look at both convenience and performance of the
libraries under consideration in the context of solving ordinary differential
equations. The comparison is based on odeint, a framework for the solution of
ordinary differential equations. We show that modern high-level libraries allow
to effectively use the computational resources of many-core GPUs or multi-core
CPUs without much knowledge of the underlying technologies. We believe the
contribution of this study warrants its publication in the SIAM Journal of
Scientific Computing.

The co-authors of the paper are: Dr. Denis Demidov (ddemidov@ksu.ru, Kazan
branch of Joint Supercomputer Center, Russian Academy of Sciences, Lobachevsky
st. 2/31, 420011 Kazan, Russia), Dr. Karsten Ahnert (karsten.ahnert@gmx.de,
Institut f\"ur Physik und Astronomie, Universit\"at Potsdam,
Karl-Liebknecht-Strasse 24/25, 14476 Potsdam-Golm, Germany), Dr. Karl Rupp
(rupp@mcs.anl.gov, Mathematics and Computer Science Division, Argonne National
Laboratory, 9700 South Cass Avenue, Argonne, IL 60439, USA), and Dr. Peter
Gottschling (Peter.Gottschling@tu-dresden.de, Technische Universit\"at Dresden,
Institut f\"ur Wissenschaftliches Rechnen, 01062 Dresden, Germany).

My coauthors and I would like to suggest, as potential referees, Dr. Nathan
Bell (wnbell@gmail.com, NVIDIA Research), Dr. Felix H\"ofling
(hoefling@mf.mpg.de, Max Planck Institute for Intelligent Systems), Dr. Andreas
Kl\"ockner (kloeckner@cims.nyu.edu, Courant Institute of Mathematical Sciences,
New York University), and Dr. Matthew Knepley (knepley@mcs.anl.gov, Computation
Institute, University of Chicago).

This paper is our original unpublished work and it has not been submitted to
any other journal for reviews. All the authors have read the paper and have
approved this submission.

\closing{Sincerely,}

\end{letter}

\end{document}
